\documentclass{article}

\usepackage{times}
\usepackage{textcomp}
\usepackage{listings}
\usepackage{fullpage}
\usepackage{color}
\usepackage{courier}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{amsmath, amsfonts, amssymb, amsthm}
\usepackage{hyperref}
\graphicspath{{./}}

\lstset{language=python, keywordstyle={\bfseries \color{green}}, basicstyle=\footnotesize\ttfamily}
\setlength{\paperheight}{11in}
\author{Clement Tsang}

\begin{document}

\begin{center}
    \Large{LS 101 --- Chapter 11: Drug Legislation and Social Policy}
\end{center}

\section{Canadian Drug Legislation}
\begin{itemize}
    \item The \emph{Controlled Drugs and Substances Act}, or CDSA, outlines prohibited drugs and penalites for posession, manufacturing, importing, and trafficking these substances.
    \item The \emph{Criminal Code} also has somewhat related laws, mainly in regards to conspiracy, anti-gang provisions, and proceeds of crime/money laundering laws.
    \item Penalties under the CDSA are severe; posession of a controlled substance can be punished by up to 7 years.
    \item Importing/exporting/producing/trafficking/conspiracy can be liable to imprisonment for life in some cases.
    \item Offences under the CDSA and \emph{Criminal Code of Canada}(CCC) include:
        \begin{enumerate}
            \item Possession of a controlled substance
            \item Double-doctoring
            \item Posession for the purpose of trafficking
            \item Trafficking
            \item Importing/exporting a controlled substance
            \item Production of a substance
            \item Posession of a property obtained by crime
            \item Laundering the proceeds of crime
            \item Conspiracy laws
            \item Participation in a criminal organization
        \end{enumerate}
\end{itemize}

\subsection{Possession}
\begin{itemize}
    \item One issue with this is what constitutes posession.
    \item The Crown must prove the accused has a guilty state of mind --- and thus, knew of the substance and its nature.
    \item It is not required for the drug to actually be what the accused thinks it is, the accused must just actually believe the drug is prohibited or is willingly blind to it being such a drug.
\end{itemize}

\subsection{Double-doctoring}
\begin{itemize}
    \item It is illegal under section 4.2 of the CDSA to seek any substances that are controlled from Schedules I to IV from a practioniner without disclosing particulars related to the acquisition of any of the scheduled substances from any other particular within the preceding 30 days.
\end{itemize}

\subsection{Posession for the purposes of trafficking, and trafficking}
\begin{itemize}
    \item Trafficking and the posession with the \emph{intent} of trafficking are both forbidden under section 5 of the CDSA.
    \item The onus is on the Crown to prove intent to traffic.
\end{itemize}

\subsection{Importing and Exporting}
\begin{itemize}
    \item It is an offence to import/export controlled substances into or out of Canada.
\end{itemize}

\end{document}
