\documentclass{article}

\usepackage{times}
\usepackage{textcomp}
\usepackage{listings}
\usepackage{fullpage}
\usepackage{color}
\usepackage{courier}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{amsmath, amsfonts, amssymb, amsthm}
\usepackage{hyperref}
\graphicspath{{./}}

\lstset{language=python, keywordstyle={\bfseries \color{green}}, basicstyle=\footnotesize\ttfamily}
\setlength{\paperheight}{11in}
\author{Clement Tsang}

\begin{document}

\begin{center}
    \Large{LS 101 --- Chapter 12: Laws and Social Policy Related to Prostitution --- Legislation of Morality}
\end{center}

\section{Prostitution Related Offences}
\begin{itemize}
    \item Prostitution is not an offence in Canada, though there are many related behaviours found in the Criminal Code.
    \item There are 3 main issues:
        \begin{enumerate}
            \item Communicating in a public place for the purposes of selling/buying sexual services.
            \item Bawdy house offences.
            \item Procuring and living off of the avails of prostitution.
        \end{enumerate}
    \item These laws mainly attempt to regulate some of the public and nuisance behaviours associated with street prostitution --- these don't really affect to higher level prostitutes.
\end{itemize}

\subsection{Communication}
\begin{itemize}
    \item Section 213 makes it an offence for a prostitute and/or the client to communicate for the purporses of prostitution.
    \item The police may attempt to enforce this by impersonating as either a prostitute or a ``John' and arrest people if they attempt to communicate for the purposes of prostitution.
    \item This is legally not entrapment if the police act on reasonable suspicion and do not actively try to lead suspects in.
    \item Studies have shown typically more women are arrested rather than male clients. 
\end{itemize}

\subsection{Bawdy-houses}
\begin{itemize}
    \item Anyone keeping a common bawdy-house is guilty of an indctable offence and liable to being imprisoned for up to two years.
\end{itemize}

\subsection{Procuring}
\begin{itemize}
    \item Anyone who:
        \begin{enumerate}
            \item Procures, attempts to procure, or solicits a person to have illicit sexual intercourse with another person, whether in/out of Canada,
            \item Entices a person who is not a prostitute to a common bawdy-house for the purpose of illicit sexual intercourse,
            \item Knowingly conceals a person in a common bawdy-house,
            \item Procurs/attempts to procure a person a prostitute,
            \item Procures or attempts to procure a preson to leave the usual place of abode of that person in Canada, with the intent that ther person may live in or frequent a common bawdy-house,
            \item Lives wholy or in part on teh avails of prostitution of another person,
        \end{enumerate}
        is guilty of an indictable offence and up to 10 years of prison.
\end{itemize}

\end{document}
